---
title: "Exploration Ideas"
format: html
---
https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-03-29/readme.md

look into rev and expend for div 1
```{r}
library(tidyverse)
library(dplyr)
# bring in the data 
tuesdata <- tidytuesdayR::tt_load('2022-03-29')
tuesdata <- tidytuesdayR::tt_load(2022, week = 13)
CollegiateSports <- tuesdata$sports

# filter so its only Division 1 sports 
div1 <- CollegiateSports |> filter(grepl('NCAA Division I', classification_name)) |> filter(classification_code <= 3)

#str detect
# check that after I, its not another I, or exactly one I

# figured out that classifications for all D1 sports are either 1, 2, or 3
div1_smaller <- div1 |> select(1,3,7,8,20:28)

# selected only the columns that I'm interested in. 

div1_rev_expend <- div1_smaller |> select(1:4, 9, 12, 13)

```


how the trend changes over time and including a graphic
- I want to make an option where you'll be able to select a school and be able to see a barplot of their total rev value and total expend value over the last 5 years 
```{r}
library(scales)
# start this with one school at a time 
# start with University of Colorado Boulder
UC_Boulder <- div1_rev_expend |> filter(institution_name == "University of Colorado Boulder") |>
  filter(!is.na(total_rev_menwomen), !is.na(total_exp_menwomen)) |>
  group_by(year) |>
  summarise(total_rev = sum(total_rev_menwomen), total_exp = sum(total_exp_menwomen))

## institution_name is set to an object, that will become input in shiny 
## use str_glue to pass object with school name through 
## use functions 


UC_Boulder2 <- UC_Boulder |> pivot_longer(cols = c(total_rev, total_exp),
                           names_to = "Type", values_to = "Dollars")

ggplot(data = UC_Boulder2, aes(x = year, y = Dollars, fill = Type)) +
  geom_col(position = "dodge") +
  scale_y_continuous("Dollars",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar()) +
  labs(title = "Total Revenue and Expenditure for Sports\nat CU Boulder 2015-2019")


# or do all schools in my roughly created D1 group
div1_collapsed <- div1_rev_expend |> filter(!is.na(total_rev_menwomen), !is.na(total_exp_menwomen)) |>
group_by(institution_name, year) |> summarise(total_rev = sum(total_rev_menwomen), total_exp = sum(total_exp_menwomen))

div1_pivot <- div1_collapsed |> pivot_longer(cols = c(total_rev, total_exp),
                           names_to = "Type", values_to = "Dollars")


## ggplot(data = div1_pivot, aes(x = year, y = Dollars, fill = Type)) +
  #geom_col(position = "dodge") +
  #facet_wrap(~institution_name) +
    #scale_y_continuous("Dollars",
    #breaks = scales::breaks_extended(8),
    #labels = scales::label_dollar())
## this is too many, try top 10 schools for rev

# top10_rev <- div1_rev_expend |> filter(!is.na(total_rev_menwomen), !is.na(total_exp_menwomen)) |>
# group_by(institution_name) |> summarise(total_rev = sum(total_rev_menwomen), total_exp = sum(total_exp_menwomen)) |>
  # arrange(desc(total_rev)) |> slice(1:10)

## create a vector with names of the universities from top 10

```

2/28 meeting with Ramler 
Goal is making a good shiny app 
find a goal of analysis
- break down by division and sport or gender 

```{r}
# figure out how to select one school using it as an object?
## I want the filter statement to be user interaction on the shiny app 
school_name <- "University of Colorado Boulder"

school_select_rev_expend <- div1_rev_expend |> filter(institution_name == school_name) |>
  filter(!is.na(total_rev_menwomen), !is.na(total_exp_menwomen)) |>
  group_by(institution_name, year) |>
  summarise(total_rev = sum(total_rev_menwomen), total_exp = sum(total_exp_menwomen)) |>
  pivot_longer(cols = c(total_rev, total_exp),
                           names_to = "Type", values_to = "Dollars")

ggplot(data = school_select_rev_expend, aes(x = year, y = Dollars, fill = Type)) +
  geom_col(position = "dodge") +
  scale_y_continuous("Dollars",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar()) +
  labs(title = str_glue("Total Revenue and Expenditure for Sports\nat {school_name}")) +
  theme_minimal() +
  scale_fill_viridis_d(option = 'cividis')
```

```{r}
div1_rev_expend |> pull(institution_name)
```


Try combining data for all schools by sport and year
  side by side boxplots 
  
Different plots for different comparisons 

Set up a tabb structure 
  tab panel 
  shiney tabs 
  
```{r}
# Now were going by sport, so go through CollegiateSports and only keep what is needed 
sport_smaller <- CollegiateSports |> select(1,3, 8, 24, 27, 28)

chose_sport <- "Football"

selected_sport <- sport_smaller |> filter(sports == chose_sport) |>
  filter(!is.na(total_rev_menwomen), !is.na(total_exp_menwomen))

# jokes, stick with only D1
# I think it might need to be narrowed down by year - it's just gonna be too much with all of these schools and more than one year 

# chose_year <- 2019

selected_sport <- div1_rev_expend |> filter(sports == chose_sport) |>
  filter(!is.na(total_rev_menwomen), !is.na(total_exp_menwomen)) |>
  group_by(institution_name, year, classification_name, sports) |>
  summarise(total_rev = sum(total_rev_menwomen), total_exp = sum(total_exp_menwomen)) |>
  pivot_longer(cols = c(total_rev, total_exp),
                           names_to = "Type", values_to = "Dollars")

## ggplot(data = selected_sport, aes(x = institution_name, y = Dollars, fill = Type)) +
  ## geom_col(position = "dodge") +
  ## scale_y_continuous("Dollars",
    ## breaks = scales::breaks_extended(8),
    ## labels = scales::label_dollar())

# Combined rev and expend for all D1 teams 
# FBS is the highest level of college football 
ggplot(data = selected_sport, aes(x = factor(year), y = Dollars, fill = Type)) +
  geom_boxplot() +
  scale_y_continuous("Dollars",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar()) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = str_glue("Total Revenue and Expenditure for\nDivision 1 {chose_sport}, 2015-2019"),
       x = "Year")

## label the school with highest rev and expend each year 
highest_rev <- 
  selected_sport |> 
  filter(Type == "total_rev") |>
  ungroup() |>
  group_by(year) |> 
  slice_max(Dollars, n = 1)

highest_expend <- 
  selected_sport |> 
  filter(Type == "total_exp") |>
  ungroup() |>
  group_by(year) |> 
  slice_max(Dollars, n = 1)

label_graph <- full_join(highest_rev, highest_expend)

library(ggrepel)
ggplot(data = selected_sport, aes(x = factor(year), y = Dollars, fill = Type)) +
  geom_boxplot() +
  geom_label(data = label_graph, aes(label = institution_name)) +
  scale_y_continuous("Dollars",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar()) 
## hard no, ugly af. Do just top 1 and then I'll do a side plot to investigate like top 5? in each 
ggplot(data = selected_sport, aes(x = factor(year), y = Dollars, fill = Type)) +
  geom_boxplot() +
  geom_label_repel(data = label_graph, aes(label = institution_name)) +
  # geom_label(data = label_graph, aes(label = institution_name)) +
  scale_y_continuous("Dollars",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar()) 

# now make a side plot showing top 5 in revenue and expend for each year 
highest_rev_5 <- 
  selected_sport |> 
  filter(Type == "total_rev") |>
  ungroup() |>
  group_by(year) |> 
  slice_max(Dollars, n = 5)

highest_expend_5 <- 
  selected_sport |> 
  filter(Type == "total_exp") |>
  ungroup() |>
  group_by(year) |> 
  slice_max(Dollars, n = 5)

top5 <- full_join(highest_rev_5, highest_expend_5)
ggplot(data = top5, aes(y = Dollars, x = factor(year))) +
  geom_point(aes(colour = Type)) +
  geom_label_repel(data = top5, aes(label = institution_name)) +
    scale_y_continuous("Dollars",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar()) 
  

```




